---
title: "ECS 132 - Homework #6"
author:  Brendan Gerrity,  \break bgerrity@ucdavis.edu # Include your name, and UC Davis email
date: "r Sys.Date()"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\pagebreak[4]

# Problem 1
Based from the proof from class, and that $Q_i = 1 - P_i$  

## 1.1
$$
  \begin{aligned}
    Q_i &= 1 - P_i \\
        &= 1 - \frac{1-(\frac{q}{p})^i}{1-(\frac{q}{p})^N}\\
        &= \frac{{1-(\frac{q}{p})^N} - {1+(\frac{q}{p})^i}}{1-(\frac{q}{p})^N} \\
        &= \frac{-(\frac{q}{p})^N + (\frac{q}{p})^i}{1-(\frac{q}{p})^N} \\
  \end{aligned}
$$
  
## 1.2
$$
  \begin{aligned}
    Q_i &= \frac{-(\frac{q}{p})^N + (\frac{q}{p})^i}{1-(\frac{q}{p})^N} \\
        &= \lim_{\frac{p}{q} \to 1} \frac{-(\frac{q}{p})^N + (\frac{q}{p})^i}{1-(\frac{q}{p})^N} \\
        &= \frac{-N + i}{-N} \\
        &= \frac{N - i}{N}
  \end{aligned}
$$

# Problem 2

$$
  \begin{aligned}
    P(i) &= \frac{1}{3}P(i+1) + \frac{2}{3}P(i-1) \\
    P(1) &= \frac{1}{3}P(2) + \frac{2}{3}P(0) \\
         &= \frac{1}{3}P(2) + 0 \\
    P(2) &= 3P(1) \\
    P(3) &= 7P(1) \\
    P(i) &= (2^i - 1)P(1) \\
    P(i+2) &= 1 = (2^{i+2} - 1)P(1) \\
    P(1)   &= \frac{1}{2^{i+2} - 1} = \frac{1}{3}P(2) \\
    P(2)   &= \frac{3}{2^{i+2} - 1} \\
           &= \frac{2^i-1}{2^{i+2} - 1} \\
    P(i)   &= \frac{2^i-1}{2^{i+2} - 1} < \frac{1}{4} \quad \forall i \geq 0\\ 
  \end{aligned}
$$
  
# Problem 3

D = 0 (Truth) | D = 1 (Lie)  
T = 0 (Pass)  | T = 1 (Fail)   
P(T = 0| D = 1) = .1  
P(T = 1| D = 0) = .2  
$$
  \begin{aligned}
    Sensitivity(n) &= 1 - P(T=1|D=0) = 0.8 \\
    Specificity(\theta) &= 1 - P(T=0|D=1) = 0.9 \\
  \end{aligned}
$$  
  
# Problem 4

1.  Plot the pdf and the cdf of $X|(D= 1)???Norm(70,15)$ for values 20 to 120.  
```{r, cache=TRUE, echo=TRUE}
par(mfrow=c(1,1))
x =seq(20,120)
hx =dnorm(x,70,15)
plot(x, hx,xlim =c(20,120),ylim =c(0,0.05),type='l',ylab="PDF X|D=1",xlab="x")
```
  
2.  Let $x*$ denote the cutoff value such that if $X > x*$ then the test is positive test otherwise it is considered be negative. Plot the ROC curve for values of $x*$ ranging from $52 < x \leq 65$.  
```{r, cache=TRUE, echo=TRUE}
par(mfrow=c(1,1))
x =seq(20,120)
hx =pnorm(x,70,15)
plot(x, hx,xlim =c(20,120),ylim =c(0,1),type='l',ylab="CDF X|D=1",xlab="x")
```
  
3.  If False Positive Rate and False Negative Rates are equally bad, determine the value of $x*$.   

# Classmate Collaborators

Inclulde the names and email IDs of everyone you collaborated with. You are free to discuss with your peers but everyone's work must be individual.
